---
title: "Ada Proposal R Code"
author: "Sulaiti Mutebi"
date: "2024-10-18"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Look at how stigma is greater than the mean (stigma \>= mean)

# Sample size being small thus becoming a limitation

# Mean of the total stigma (continous variable) and depression (categorical yes/no)

# Total stigma as the continuous variable

# Test for the assumption

# See the difference in the box plot

# Variables of interest:

#1. AGE: (1=50-54, 2=55-59, 3=60+) 

#2. SEX: (1=Female, 2= Male) 

#3.ETHNICS: (0=White, 1= Non-White)

#4.TOTSTIG (stigma)

#5.TOTCESD(depression)

#Stigma (1= mean \<=, 2= mean \>=)\~independent variable 

#Depression (1=yes, 0=no)\~dependent varaiable


```{r}

# Load necessary libraries
library(dplyr)
library(ggplot2)
pacman::p_load(ggplot2, car, odds.n.ends, readr, dplyr)
library(table1)
suppressWarnings(library(table1))
library(gmodels)
```

```{r}

# loading the dataset

load("33242-0001-Data.rda") # Ensure the .rda file is in the working directory
ProposalDataset <- da33242.0001

# Summary of dataset to check the variables
summary(ProposalDataset)

```
#### Select variables of interest and renaming them

``````{r}
# Select variables of interest and rename them
Dataset <- ProposalDataset %>%
  select(AGE, SEX, ETHNIC2, TOTSTIG, TOTCESD) %>%
  rename(
    age_category = AGE,
    sex = SEX,
    race = ETHNIC2,
    stigma = TOTSTIG,
    depression = TOTCESD

  )
```
  
#### Data management for the analysis
 
```{r}
# Reclassify depression into binary categories
Dataset <- Dataset %>%
  mutate(
    depression_binary = case_when(
      depression >= 0 & depression < 16 ~ 0, # Not Depressed
      depression >= 16 ~ 1                # Depressed
    )
  )

# Convert depression_binary to a factor
Dataset$depression_binary <- factor(
  Dataset$depression_binary,
  levels = c(0, 1),
  labels = c("No", "Yes")
)
```
 
#### Recoding and Factoring Other Variables
```{r}
# Recode and factor variables
Dataset <- Dataset %>%
  mutate(
    age_category = factor(age_category, levels = 1:3, labels = c("50-54 years old", "55-59 years old", "60+ years")),
    sex = factor(sex, levels = 1:2, labels = c("Female", "Male")),
    race = factor(race, levels = 0:1, labels = c("White", "Non-White"))
  )
```

### Data Verification
```{r}
# Check the structure and summary of the Dataset
str(Dataset)
summary(Dataset)

# Cross-tabulation of the depression_binary variable
table(Dataset$depression_binary, useNA = "always")
```

### Visualization: Boxplot of Stigma by Depression Status

```{r}
# Boxplot of stigma by depression status
ggplot(Dataset, aes(x = depression_binary, y = stigma, fill = depression_binary)) +
  geom_boxplot() +
  labs(
    title = "Boxplot of Stigma by Depression Status",
    x = "Depression Status (Yes/No)",
    y = "Total Stigma Score"
  ) +
  theme_minimal()
```

### Drawing my dag diagram

```{r}
library(dagitty)

dg <- dagitty('
  dag {
    
    race-> stigma
    Sex -> depression_binary
    Sex -> stigma
   age -> depression_binary
   stigma -> depression_binary
    
  }
')

# Plot the DAG
plot(dg)

```

### Running a univariate logistic regression model for stigma and ORs and 95% CIs.

```{r}

# A univariate logistic regression model
model <- glm(depression_binary ~ stigma, 
             data = Dataset, 
             family = binomial)

# Summary of the model
summary(model)

odds.n.ends(model)

```

### Table1 not stratified

```{r}

table1(~ age_category+sex+race+stigma+depression_binary,data = Dataset)

```

### Table one stratified by stigma

```{r}
#creating table 1 and data summary
table1(~ age_category+sex+race+depression_binary|stigma,data = Dataset)

```

### Table one stratified by depression

```{r}
#creating table 1 and data summary
table1(~ age_category+sex+race+stigma|depression_binary,data = Dataset)

```

### Running a multivariate logistic regression model that includes stigma, age_category,race, and sex in the model as covariates. 

```{r}
model1 <- glm(depression_binary ~ age_category+sex+race+stigma, data = Dataset, family = binomial)
family = binomial(link = "logit")

summary(model1)

odds.n.ends(model)
```

### fiting my model using broom tidy to get estimates

```{r}

# Calculate Odds Ratios (OR) and 95% Confidence Intervals (CI)
broom::tidy(model1, exponentiate = TRUE, conf.int = TRUE)
```

### Model diagnostics: check multicollinearity using Variance Inflation Factor (VIF)

```{r}
car::vif(model1)
```

### Check interaction effect between HIV stigma and race

```{r}

# Fit a model with interaction term
model_interaction <- glm(depression_binary ~ stigma * race, 
                         data = Dataset, 
                         family = binomial)

# Summary of the interaction model
summary(model_interaction)

```

### Check interaction effect between HIV stigma and age_category

```{r}

# Fit a model with interaction term
model_interaction1 <- glm(depression_binary ~ stigma * age_category, 
                         data = Dataset, 
                         family = binomial)

# Summary of the interaction model
summary(model_interaction1)


```

### Check interaction effect between HIV stigma and sex

```{r}

# Fit a model with interaction term
model_interaction2 <- glm(depression_binary ~ stigma * sex, 
                         data = Dataset, 
                         family = binomial)

# Summary of the interaction model
summary(model_interaction2)


```

### Compare the models (with and without interactions)

```{r}

# Compare models using ANOVA
anova(model, model_interaction,model_interaction1,model_interaction2, test = "Chisq")

```

### Visualize the results-predicted probabilities

```{r}

# Generate predicted probabilities
Dataset$predicted_probs <- predict(model, type = "response")

# Plot predicted probabilities by stigma
ggplot(Dataset, aes(x = stigma, y = predicted_probs, color = depression_binary)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "loess", se = FALSE) +
  labs(
    title = "Predicted Probabilities of Depression by Stigma",
    x = "Total Stigma Score",
    y = "Predicted Probability of Depression"
  ) +
  theme_minimal()


```

### Visualize the results-interaction effect

```{r}
# Visualize interaction effect
ggplot(Dataset, aes(x = stigma, y = predicted_probs, color = race)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "loess", se = FALSE) +
  facet_wrap(~ depression_binary) +
  labs(
    title = "Interaction Between Stigma and Race on Depression",
    x = "Total Stigma Score",
    y = "Predicted Probability of Depression"
  ) +
  theme_minimal()

```








