---
title: "Ada Proposal R Code"
author: "Sulaiti Mutebi"
date: "2024-10-18"
output:
  word_document: default
  html_document: default
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Sample size being small thus becoming a limitation

#Total stigma as the continuous variable

#Test for the assumption

#See the difference in the box plot

#Variables of interest:

#1. AGE: (1=50-54, 2=55-59, 3=60+)

#2. SEX: (1=Female, 2= Male)

#3.ETHNICS: (0=White, 1= Non-White)

#4.TOTSTIG (stigma)

#5.TOTCESD(depression)

#Stigma (continuous)\~independent variable

#Depression (1=yes, 0=no)\~dependent varaiable

```{r}

# Load necessary libraries
library(dplyr)
library(ggplot2)
library(car)
library(odds.n.ends)
library(readr)
library(table1)
suppressWarnings(library(pacman))
pacman::p_load(ggplot2, gmodels, broom)
```

```{r}

# loading the dataset

load("33242-0001-Data.rda") 
ProposalDataset <- da33242.0001

# Summary of dataset to check the variables
summary(ProposalDataset)

```

#### Select variables of interest and renaming them

```{r}
# Select variables of interest and rename them
Dataset <- ProposalDataset %>%
  select(AGE, SEX, ETHNIC2, TOTSTIG, TOTCESD) %>%
  rename(
    age_category = AGE,
    sex = SEX,
    race = ETHNIC2,
    stigma = TOTSTIG,
    depression = TOTCESD

  )
```

#### Data management for the analysis

```{r}
# Reclassify depression into binary categories
Dataset <- Dataset %>%
  mutate(
    depression_binary = case_when(
      depression >= 0 & depression < 16 ~ 0, # Not Depressed
      depression >= 16 ~ 1                # Depressed
    )
  )

# Convert depression_binary to a factor
Dataset$depression_binary <- factor(
  Dataset$depression_binary,
  levels = c(0, 1),
  labels = c("No", "Yes")
)
```

### Data Verification

```{r}
# Check the structure and summary of the Dataset
str(Dataset)
summary(Dataset)

# Cross-tabulation of the depression_binary variable
table(Dataset$depression_binary, useNA = "always")
```

### Visualization: Boxplot of Stigma by Depression Status

```{r}
# Boxplot of stigma by depression status
ggplot(Dataset, aes(x = depression_binary, y = stigma, fill = depression_binary)) +
  geom_boxplot() +
  labs(
    title = "Boxplot of Stigma by Depression Status",
    x = "Depression Status (Yes/No)",
    y = "Total Stigma Score"
  ) +
  theme_minimal()
```

### Linearity Assumption Test for Stigma

```{r}

# Fit a logistic regression model
log_model <- glm(depression_binary ~ stigma, data = Dataset, family = binomial)

# Create a binned residual plot to assess linearity
Dataset <- Dataset %>%
  mutate(stigma_bin = cut(stigma, breaks = 10))

binned_means <- Dataset %>%
  group_by(stigma_bin) %>%
  summarize(mean_stigma = mean(stigma, na.rm = TRUE),
            mean_residual = mean(residuals(log_model, type = "response"), na.rm = TRUE))

# Plot residuals to check linearity
ggplot(binned_means, aes(x = mean_stigma, y = mean_residual)) +
  geom_point() +
  geom_smooth(method = "loess") +
  labs(
    title = "Linearity Assumption Check",
    x = "Mean Stigma (Binned)",
    y = "Mean Residual"
  ) +
  theme_minimal()

# Summary of findings:
```

#Interpretation of the linear assumption output 
The flat line and near-zero residuals indicate that the linearity assumption for the
stigma variable is satisfied in the logistic regression model. No
evidence of non-linearity is observed.

### Drawing my dag diagram

```{r}
library(dagitty)

dg <- dagitty('
  dag {
    
    race-> stigma
    Sex -> depression_binary
    Sex -> stigma
   age -> depression_binary
   stigma -> depression_binary
    
  }
')

# Plot the DAG
plot(dg)

```

### Running a univariate logistic regression model for stigma and ORs and 95% CIs.

```{r}

# A univariate logistic regression model
model <- glm(depression_binary ~ stigma, 
             data = Dataset, 
             family = binomial)

# Summary of the model
summary(model)

odds.n.ends(model)

Dataset$stigma

```

### Table1 not stratified

```{r}

table1(~ age_category+sex+race+stigma+depression_binary,data = Dataset)

```

### Table one stratified by stigma

```{r}
#creating table 1 and data summary
table1(~ age_category+sex+race+depression_binary|stigma,data = Dataset)

```

### Table one stratified by depression

```{r}
#creating table 1 and data summary
table1(~ age_category+sex+race+stigma|depression_binary,data = Dataset)

```

### Running a multivariate logistic regression model that includes stigma, age_category,race, and sex in the model as covariates.

```{r}
model1 <- glm(depression_binary ~ age_category+sex+race+stigma, data = Dataset, family = binomial)
family = binomial(link = "logit")

summary(model1)

odds.n.ends(model)

```

### fiting my model using broom tidy to get estimates

```{r}

# Calculate Odds Ratios (OR) and 95% Confidence Intervals (CI)
broom::tidy(model, exponentiate = TRUE, conf.int = TRUE)
```

### Model diagnostics: check multicollinearity using Variance Inflation Factor (VIF)

```{r}
car::vif(model1)
```

### Check interaction effect between HIV stigma and race

```{r}

# Fit a model with interaction term
model_interaction <- glm(depression_binary ~ stigma * race, 
                         data = Dataset, 
                         family = binomial)

# Summary of the interaction model
summary(model_interaction)

```

### Check interaction effect between HIV stigma and age_category

```{r}

# Fit a model with interaction term
model_interaction1 <- glm(depression_binary ~ stigma * age_category, 
                         data = Dataset, 
                         family = binomial)

# Summary of the interaction model
summary(model_interaction1)


```

### Check interaction effect between HIV stigma and sex

```{r}

# Fit a model with interaction term
model_interaction2 <- glm(depression_binary ~ stigma * sex, 
                         data = Dataset, 
                         family = binomial)

# Summary of the interaction model
summary(model_interaction2)


```

### Compare the models (with and without interactions)

```{r}

# Compare models using ANOVA
anova(model, model_interaction,model_interaction1,model_interaction2, test = "Chisq")

```

### Visualize the results-predicted probabilities

```{r}

# Generate predicted probabilities
Dataset$predicted_probs <- predict(model, type = "response")

# Plot predicted probabilities by stigma
ggplot(Dataset, aes(x = stigma, y = predicted_probs, color = depression_binary)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "loess", se = FALSE) +
  labs(
    title = "Predicted Probabilities of Depression by Stigma",
    x = "Total Stigma Score",
    y = "Predicted Probability of Depression"
  ) +
  theme_minimal()


```

### Visualize the results-interaction effect

```{r}
# Visualize interaction effect
ggplot(Dataset, aes(x = stigma, y = predicted_probs, color = race)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "loess", se = FALSE) +
  facet_wrap(~ depression_binary) +
  labs(
    title = "Interaction Between Stigma and Race on Depression",
    x = "Total Stigma Score",
    y = "Predicted Probability of Depression"
  ) +
  theme_minimal()

```
